Measurement and Probability in Fuzzy Quantum Theories

### 2016-11-25 proposal/prop.tex
prop

\item If we insist that probabilities cannot be infinitely precise and
  are bound to be approximations represented by intervals of
  confidence, then by Gleason's theorem, quantum states themselves
  cannot be infinitely precise. This raises an important philosophical
  question with immense practical relevance in the context of quantum
  computing. The question, which we could largely bypass in the
  classical case, is whether there is, in the background of all the
  measurements and probabilities, a ``real'' infinitely-precise
  quantum state that is being approximated and that, in principle, can
  be more and more accurately described given more and more
  resources. The answer will have to be \emph{no} if we are to respect
  major theorems of quantum mechanics and quantum information. Thus
  our first radical departure from the classical case will be that the
  quantum state itself must be fuzzy in some sense to be
  formalized. In other words, does a Gleason, Bell, or Kochen-Specker
  theorem still hold in a world with finite resources?

\item Going further we actually have to concede not just the infinite
  accuracy of the quantum state, but its very existence as a ``real''
  entity that exists independently of measurements and
  probabilities. The reasons for this radical position are as
  follows. First arguments embedded in the proof of the Kochen-Specker
  theorem essentially rule out the existence of a sample space as the
  probabilities of events cannot be consistently expressed using core
  assignments to elements of the sample space. Second, the frequency
  interpretation of probability which is already problematic in the
  classical case, becomes untenable in the quantum case. In both
  cases, but much more in the quantum case, it is much more sensible
  to interpret probability in the Bayesian sense as ``lack of
  knowledge''. This, and similar arguments elegantly expressed by
  recent work on Quantum Bayesianism~\cite{Fuchs2010,VonBaeyer2016},
  suggest that the quantum state is more like an interactive system in
  computer science parlance. In other words, the state is only
  determined by its interactions with the environment or observers. A
  rough but useful analogy in the context of computer science would be
  a web server: the state of such a system cannot be said to exist
  independently of the interactions with clients; each interaction
  with a client determines a future state of the system and may affect
  future interactions. Crucially, each client may have a
  \emph{different} view of the web server that is consistent with
  their own history of interactions. This computational perspective
  agrees with the philosophical position of QBism~\cite{VonBaeyer2016}
  that asserts that the quantum state is subjective: each observer has
  a different view of the quantum system that is consistent with their
  previous observations and that allows that observer, independently
  of other observers, to assign beliefs (i.e., probabilities) to
  future interactions.

@@

The early formalization of the process of computing using Turing
machines~\cite{citeulike:321509} or the
$\lambda$-calculus~\cite{Barendregt:Lambda} viewed a computation as
realizing a mathematical function. From our perspective, the most
important consequence of this view is that the properties of the
computed function existed independently of any observation,
calculation, or interaction. An agent can only interact with the
computation by asking questions regarding its input-output behavior
and although such interaction would reveal fragments of the computed
function it could not change it or affect it in any way.

In a series of breakthrough papers that led to his Turing
Award~\cite{Milner:1993:EIT:151233.151240}, Milner argued that this
``closed system'' view of computation is too limiting as it cannot
adequately describe \emph{interactive computations}. Milner explained
that ``concurrency requires a new conceptual framework'' and ``not
merely an extension of the repertoire of entities and constructions
which explain sequential computing.'' He gives the following simple
but compelling example. Consider the following two programs:
\begin{verbatim}
Program P1: x := 1 ; x := x + 1 
Program P2: x := 2
\end{verbatim}
As closed systems, these two programs define the same relation, namely
the one that maps an input memory to an updated memory in which the
location named \verb|x| is assigned the value \verb|2|. But place the
programs in an environment which may involve interactions and the
equivalence is lost. Concretely, place each of the two programs
concurrently with the program:
\begin{verbatim}
Program Q: x := 3
\end{verbatim}
and we can observe the following. In the combined system of \verb|P1|
and \verb|Q| running concurrently, the location \verb|x| may end up
containing \verb|2|, \verb|3|, or \verb|4|. But in the combined system
of \verb|P2| and \verb|Q| running concurrently, the location~\verb|x|
can only end up containing \verb|2| or \verb|3|. The lesson, phrased
in physics terms, is that it is not sensible to think of the properties
of a system (like \verb|P1| or \verb|P2|) independently of the
observer (\verb|Q| in this case). Indeed the observation \verb|x=4|
cannot said to have an independent existence independently from the
interaction between \verb|P1| and \verb|Q|: it really emerges or is
created as part of the observation process and is not a pre-existing
``element of reality''.

@@

do quantum version; application to CS and quantum information

>>>>

\item If we insist that probabilities cannot be infinitely precise and
  are bound to be approximations represented by intervals of
  confidence, then by Gleason's theorem, quantum states themselves
  cannot be infinitely precise. This raises an important philosophical
  question with immense practical relevance in the context of quantum
  computing. The question, which we could largely bypass in the
  classical case, is whether there is, in the background of all the
  measurements and probabilities, a ``real'' infinitely-precise
  quantum state that is being approximated and that, in principle, can
  be more and more accurately described given more and more
  resources. In other words, does a Gleason, Bell, or Kochen-Specker
  theorem still hold in a world with finite resources?  The answer
  will have to be \emph{no} if we are to respect these major theorems
  of quantum mechanics and quantum information. Thus our first radical
  departure from the conventional case will be that the quantum state
  itself must be fuzzy in some sense to be formalized.

\item Going further we actually have to concede not just the infinite
  accuracy of the quantum state, but its very existence as a ``real''
  entity that exists independently of measurements and
  probabilities. The reasons for this second radical position are as
  follows. First arguments embedded in the proof of the Kochen-Specker
  theorem essentially rule out the existence of a sample space as the
  probabilities of events cannot be consistently expressed using core
  assignments to elements of the sample space. Second, the frequency
  interpretation of probability which is already problematic in the
  classical case, becomes untenable in the quantum case. In both
  cases, but much more in the quantum case, it is much more sensible
  to interpret probability in the Bayesian sense as ``lack of
  knowledge''. This, and similar arguments elegantly expressed by
  recent work on Quantum Bayesianism~\cite{Fuchs2010,VonBaeyer2016},
  suggest that the quantum state is more like an interactive system in
  computer science parlance. In other words, the state is only
  determined by its interactions with the environment or observers. A
  rough but useful analogy in the context of computer science would be
  a web server: the state of such a system cannot be said to exist
  independently of the interactions with clients; each interaction
  with a client determines a future state of the system and may affect
  future interactions. Crucially, each client may have a
  \emph{different} view of the web server that is consistent with
  their own history of interactions. This computational perspective
  agrees with the philosophical position of QBism~\cite{VonBaeyer2016}
  that asserts that the quantum state is subjective: each observer has
  a different view of the quantum system that is consistent with their
  previous observations and that allows that observer, independently
  of other observers, to assign beliefs (i.e., probabilities) to
  possible future interactions.

@@

The early formalization of the process of computing using Turing
machines~\cite{citeulike:321509} or the
$\lambda$-calculus~\cite{Barendregt:Lambda} viewed a computation as
realizing a mathematical function. From our perspective, the most
important consequence of this view is that the properties of the
computed function existed independently of any observation,
calculation, or interaction. For example, a function is
\emph{injective} or not independently of any observation. An agent can
only interact with the computation by asking questions regarding its
input-output behavior and although such interaction would reveal
fragments of the computed function and its innate properties it could
not change it or affect it in any way.

In a series of breakthrough papers that led to his Turing
Award~\cite{Milner:1993:EIT:151233.151240}, Milner argued that this
``closed system'' view of computation is too limiting as it cannot
adequately describe \emph{interactive computations}. Milner explained
that ``concurrency requires a new conceptual framework'' and ``not
merely an extension of the repertoire of entities and constructions
which explain sequential computing.'' In his Turing Award lecture, he
gives the following simple but compelling example. Consider the
following two programs:
\begin{verbatim}
Program P1: x := 1 ; x := x + 1 
Program P2: x := 2
\end{verbatim}
As closed systems, these two programs define the same relation, namely
the one that maps an input memory to an updated memory in which the
location named \verb|x| is assigned the value \verb|2|. An observer
with the ability of inquire about the value of \verb|x| at the end of
the computation cannot distinguish the two programs. But place the
programs in an environment which may involve interactions and the
equivalence is lost. Phrased in physical terms, give the observer the
additional ability to access the location \verb|x| \emph{during} the
computation, not just to observe its final value. In computing terms,
we place each of the two programs \verb|P1| and \verb|P2| concurrently
with the program:
\begin{verbatim}
Program Q: x := 3
\end{verbatim}
In the combined system of \verb|P1| and \verb|Q| running concurrently,
the location \verb|x| may end up containing \verb|2|, \verb|3|, or
\verb|4|. But in the combined system of \verb|P2| and \verb|Q| running
concurrently, the location~\verb|x| can only end up containing
\verb|2| or \verb|3|. The lesson, phrased in physics terms, is that it
is not sensible to think of the properties of a system (like \verb|P1|
or \verb|P2|) independently of the observer (\verb|Q| in this
case). Indeed the observation \verb|x=4| cannot said to have an
existence independently from the interaction between \verb|P1| and
\verb|Q|: it really emerges or is created as part of the observation
process and is not a pre-existing ``element of reality''.

@@

\paragraph*{Research Question.} Is it possible to adapt the theory of
observation of concurrent processes to the study of quantum states and
confirm that the observations exhibit the same Bell-like correlations?

\paragraph*{Research Question.} The theory of observation of
concurrent processes and the theory of measurement in quantum
mechanics are superficially different. Is it possible to reformulate
them in a common language to benefit each theory from insights in the
other?
